Discrete Quantum Theories and Computing

### 2019-02-09 dissertation.tex
dissertation.lyx Acknowledgments Updates...

\noindent \centering{}\caption{\label{tab:committee}I want to thank the professors co-authored with
me, suggested me as a committee member, and taught me; the collaboration
or supervised works or studies become the basis of or inspire this
dissertation.}

@@

 & Co-author & Co-author & Co-author &  &  &  & Most of Chapters~\ref{chap:Conventional-Quantum-Theory}, \ref{chap:Unrestricted-Finite-Fields},and~\ref{chap:DQT=000026DQC}
is based on~\cite{geometry2013} and~\cite{DQT2014}; most of Chapter~\ref{chap:QIVPM}
is based on~\cite{THOS2018}.\tabularnewline

@@

Research committee & Co-chair & Co-chair & Member & Co-chair & Member &  & Some questions in the dissertation proposal are answered in Chapter~\ref{chap:QIVPM},and
some of the remaining ones are explained in Chapter~\ref{chap:Further-Discussion}.\tabularnewline

@@

Tier 3 committee & Co-chair & Co-chair &  & Member &  & Co-chair & Two of the assigned papers in the Tier 3 exam \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644}
are cited in Sec.~\ref{modalquantum}.\tabularnewline

@@

To have the current the content of the dissertation, I would also
like to thank John Gardiner for some inspiring discussion about quantum
probability measures over finite fields~\cite{Gardiner2014}. His
results were improved and motivated the transition from DQT to QIVPM
in Sec.~\ref{sec:Toward-IVPM}. Besides of my committee members,
the knowledge I learned from teachers in other courses help me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Although
not directly relating to the dissertation, I have to thank for your
teaching to make me more mature and easier to communicate with others,
especially my English teachers, Traci Nagle, Elizabeth ``Betsy''
Merceron, and Kexin ``Casey'' Chen (谌可心).
\begin{table}[t]
\begin{doublespace}
\noindent \centering{}\caption{\label{tab:teacher}The knowledge I learned from teachers in other
courses help me consolidate this dissertation, where IUB stands for
Indiana University Bloomington and NTU stands for National Taiwan
University.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{Jin-Tzu~Chen}cM{MATH-M522}Y}
\toprule 
Instructor & School & Course & Relation with the dissertation\tabularnewline
\midrule
\midrule 
Tom Lewis & IUB & SLST-T501 & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper of \emph{Academic Writing}.\tabularnewline
\midrule 
Michael\nolinebreak[3] A. Mandell & IUB & MATH-M522 & The textbook~\cite{Hatcher2001} of \emph{Topology II} is cited
in Chapters~\ref{chap:Introduction} and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline
\midrule 
Steven Myers & IUB & CSCI-B502 & The textbook~\cite{AroraBarak2009} of \emph{Computational Complexity}
is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline
\midrule 
Valery Lunts & IUB & MATH-M501 & The textbook~\cite{DummitFoote2004} of \emph{Survey of Algebra}
is cited in Sec.~\ref{sec:background}, \ref{discretequantumtheoryI},
and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline
\midrule 
Huah Chu\linebreak[3] (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & The textbook~\cite{Artin1991} of \emph{Algebra (I)} and \emph{(II)}
is cited in Sec.~\ref{CQC1qubitBloch.sec}, \ref{sec:background},
\ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline
\midrule
Jin-Tzu Chen\linebreak[3] (陳金次) & NTU & 201~31300 & The textbook~\cite{GAMELIN2003} of \emph{Functions of a Complex Variable}
is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{doublespace}
\end{table}

Many people in both the departments of Mathematics and Computer Science
have done everything to make my double-major smooth. \todo{Acknowledgments
cleaned up until here. }I need to thank all of them, especially directors
of graduate study in Mathematics, especially Prof.\ Christopher Judge,
Prof.\ Matthias Weber, and Prof.\ Michael A.\ Mandell; directors
of Ph.D. study in Computer Science, especially Prof.\ Yuqing Melanie
Wu (吴愈青) and Prof.\ \textturkish{Funda Ergun}.

@@

\subsection{Maximal entanglement}

@@

Performing the computation using equation (\ref{eq:john}), we find
the numbers of maximally entangled states for two qubits to be 
\begin{equation}
p\left(p^{2}-1\right)\left(p+1\right)=\left\{ 96,2688,15\,840,136\,800,\ldots\right\} \,.
\end{equation}
The irreducible state counts for maximal entanglement are reduced
by $\left(p+1\right)$, giving, for $n=2$, 
\begin{equation}
p\left(p^{2}-1\right)=\left\{ 24,336,1320,6840,\ldots\right\} \,.
\end{equation}
For three qubits, there are $p^{3}\left(p^{4}-1\right)\left(p+1\right)$
(total) and $p^{3}\left(p^{4}-1\right)$ (irreducible) instances of
pure maximally entangled states, while the general formula for 4-qubit
states remains unclear.

>>>>

\noindent \centering{}\caption{\label{tab:committee}The professors served in all my committees.}

@@

 & Co-author & Co-author & Co-author &  &  &  & Most of Chapters~\ref{chap:Conventional-Quantum-Theory}, \ref{chap:Unrestricted-Finite-Fields},
and~\ref{chap:DQT=000026DQC} is based on~\cite{geometry2013} and~\cite{DQT2014};
most of Chapter~\ref{chap:QIVPM} is based on~\cite{THOS2018}.\tabularnewline

@@

Research committee & Co-chair & Co-chair & Member & Co-chair & Member &  & Some questions in the dissertation proposal are answered in Chapter~\ref{chap:QIVPM},
and some of the remaining ones are explained in Chapter~\ref{chap:Further-Discussion}.\tabularnewline

@@

Tier 3 committee & Co-chair & Co-chair &  & Member &  & Co-chair & Two of the assigned papers \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644}
in the Tier 3 exam~\cite{Mathematics2018} are cited in Sec.~\ref{modalquantum}.\tabularnewline

@@

To have the current the content of the dissertation, I would also
like to thank John Gardiner for some inspiring discussion about quantum
probability measures over finite fields~\cite{Gardiner2014}. His
research supervised by Prof.\ Gerardo Ortiz inspired Sec.~\ref{subsec:Maximal-entanglement}
and~\ref{sec:Toward-IVPM}. Besides of my committee members, the
knowledge I learned from teachers in other courses help me consolidate
this dissertation as listed in Table~\ref{tab:teacher}. Although
not directly relating to the dissertation, I have to thank for your
teaching to make me more mature and easier to communicate with others,
especially my English teachers, Traci Nagle, Elizabeth ``Betsy''
Merceron, and Kexin ``Casey'' Chen (谌可心).
\begin{table}[t]
\begin{doublespace}
\noindent \centering{}\caption{\label{tab:teacher}The knowledge I learned from teachers in other
courses help me consolidate this dissertation, where IUB stands for
Indiana University Bloomington and NTU stands for National Taiwan
University.}
%% ERT block 1 (before LyX tabular)
\renewcommand{\tabular}{\tabularx{\linewidth}}
\renewcommand{\endtabular}{\endtabularx} 
\begin{tabular}{M{Jin-Tzu~Chen}cM{MATH-M522}Y}
\toprule 
Instructor & School & Course & Relation with the dissertation\tabularnewline
\midrule
\midrule 
Tom Lewis & IUB & SLST-T501 & Sec.~\ref{subsec:Discrete-Deutsch-algorithm} is based on the term
paper of \emph{Academic Writing}.\tabularnewline
\midrule 
Michael\nolinebreak[3] A. Mandell & IUB & MATH-M522 & The textbook~\cite{Hatcher2001} of \emph{Topology II} is cited
in Chapters~\ref{chap:Introduction} and~\ref{chap:Conventional-Quantum-Theory}.\tabularnewline
\midrule 
Steven Myers & IUB & CSCI-B502 & The textbook~\cite{AroraBarak2009} of \emph{Computational Complexity}
is cited in Sec.~\ref{modalquantumcomputing}.\tabularnewline
\midrule 
Valery Lunts & IUB & MATH-M501 & The textbook~\cite{DummitFoote2004} of \emph{Survey of Algebra}
is cited in Sec.~\ref{sec:background}, \ref{discretequantumtheoryI},
and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline
\midrule 
Huah Chu\linebreak[3] (朱樺\textenglish[variant=american]{)} & NTU & 221~U3830\linebreak[3] 221~U3840 & The textbook~\cite{Artin1991} of \emph{Algebra (I)} and \emph{(II)}
is cited in Sec.~\ref{CQC1qubitBloch.sec}, \ref{sec:background},
\ref{discretequantumtheoryI}, and~\ref{sec:Discrete-Quantum-Theory-(II)}.\tabularnewline
\midrule
Jin-Tzu Chen\linebreak[3] (陳金次) & NTU & 201~31300 & The textbook~\cite{GAMELIN2003} of \emph{Functions of a Complex Variable}
is cited in Sec.~\ref{subsec:Vector-Spaces}.\tabularnewline
\bottomrule
\end{tabular}% ERT block 2 (after LyX tabular)
\renewcommand{\tabular}{\ORIGtabular}
\renewcommand{\endtabular}{\ORIGendtabular} 
\end{doublespace}
\end{table}

Many people have done everything to make my double-major smooth and
to financially support me during my Ph.D. study. I need to thank all
of them, and the complete list of the support according to the employment
and award is listed in my Curriculum Vitae at the end of the dissertation.
Most importantly Prof.\ Amr A. Sabry\ (\textarabic{عمرو صبري})
helped me deal with the rules, supervised my reading courses for writing
our papers, and financially support me as a research assistant and
associate. In the department of Mathematics, I need to thank the graduate
services assistant, Kate Forrest, and the directors of graduate study
in Mathematics, especially Prof.\ Christopher Judge, Prof.\ Matthias
Weber, and Prof.\ Michael A.\ Mandell. In the department of Computer
Science, I want to thank the graduate student office; the director
of AI/UI assignments, Charles Pope; the directors of Ph.D. study in
Computer Science, especially Prof.\ Yuqing Melanie Wu (吴愈青) and Prof.\ \textturkish{Funda
Ergun}; the infrastructure and technology group (ITG). In particular,
I don't know how much computational power I consumed to count the
states in Sec.~\ref{subsec:Maximal-entanglement}. These computational
power not only came from the Linux systems in Computer Science but
also came from the Big Red II administrated by the University Information
Technology Services (UITS). Also, I have checked out a stack of books
from the Indiana University Libraries. While the books are not in
the collection, I occasionally requested them via the Interlibrary
Loan service, submitted the Purchase Recommendation Form to suggested
IU libraries to buy them.\todo{Acknowledgments cleaned up until here.
}

Except the financial support via employment and award, I have also
support by my family, especially my father Cheng-Tien Tai (戴振沺).

@@

\subsection{Maximal entanglement\label{subsec:Maximal-entanglement}}

@@

Performing the computation using equation (\ref{eq:john}), we find
the numbers of maximally entangled states for two qubits to be 
\begin{equation}
p\left(p^{2}-1\right)\left(p+1\right)=\left\{ 96,2688,15\,840,136\,800,\ldots\right\} \,.
\end{equation}
The irreducible state counts for maximal entanglement are reduced
by $\left(p+1\right)$, giving, for $n=2$, 
\begin{equation}
p\left(p^{2}-1\right)=\left\{ 24,336,1320,6840,\ldots\right\} \,.
\end{equation}
For three qubits, there are $p^{3}\left(p^{4}-1\right)\left(p+1\right)$
(total) and $p^{3}\left(p^{4}-1\right)$ (irreducible) instances of
pure maximally entangled states. For four-qubits and $p=3$, there
are $2195538048$ instances of pure maximally entangled states, while
the general formula for four-qubit states remains unclear.
